SAT Competition 2020\tnoteref{title}

### 2021-06-28 paper/main.tex
conclusion spellcheck

For data-driven selection of benchmark instances, we used GBD~Tools\footnote{\url{See https://pypi.org/project/gbd-tools/}} which facilitates querying for instances with desired properties, e.g., by instance author, family, result or solver runtime~\cite{Iser:2018:GBD}. 
We also use GBD~Tools for distributing benchmark instances and their attributes to the general public.\footnote{See \url{https://gbd.iti.kit.edu}}

@@

The ``Bring Your Own Benchmarks'' (BYOB) rule, first  established in SAT Competition 2017~\cite{SC2017}, was again followed
in 2020.
By this rule, solver authors are required to submit 20 ``new'' benchmark instances to accompany a solver submission in order to participate in the competition. 
New refers to the fact that instances contained in benchmark sets from earlier SAT Competitions are not allowed.
Furthermore, a least 10 of the required 20 instances are required to be ``interesting'', interpreted in loose terms by
 the standard \solver{Minisat} SAT solving needing at least one minute of runtime (on typical computing hardware) to solve an instance.
It should be noted that new benchmarks could be submitted to the competition without needing to submit a solver.
As a results, as detailed in Table~\ref{tab:families}, 27 authors contributed a set of 1260 new benchmark instances from a wide range of different instance families.

@@

To compile the set of 300 new instances, we first applied a hardness criterion by filtering out all instances solved by \solver{Minisat} in less than 10 minutes.\footnote{We note that
the limit of 10 minutes is again somewhat arbitrary.
This 10-min runtime hardness filter essentially aims to make sure that enough instances are included in the final benchmark set which allow for distinguishing in terms of relative performance between the best-performing competing solvers.
Unfortunately this time the 10-min limit enforced  was much greater that the requirements imposed for  ``interesting'' benchmark instances
by the BYOB rule. It could be more sensible to impose the some 10-min limited also for an instance being ``interesting''. However, this would require more efforts
at least in terms of computation times from the solver authors in order to construct a set of 20 interesting new benchmarks required by the BYOB rule.}

From the resulting 1012 instances, in order to obtain a balanced benchmark set, 
we randomly selected $k$~instances per author using the value  $k$ which ensured that  the resulting set contains at least 300 instances.
This \emph{rule-based randomization} procedure is detailed as
Algorithm~\ref{algo:select}. Specifically,
we randomly selected 7~satisfiable and 7~unsatisfiable instances per author (Lines~3 and~4) and added instances of yet unknown result if this did not yield a total of 14~instances (Lines~5--7). 
Of the such obtained 308~instances, we randomly removed 8~satisfiable instances, yielding a total of 114~satisfiable, 78~unsatisfiable and 108~instances of unknown satisfiability
status.

@@

For the Planning track,
the benchmark instances were generated using two SAT-based planners
\emph{Madagascar}~\cite{Madagascar14} and \emph{Pasar}~\cite{Pasar19}. We used
Madagascar both in its default configuration to generate a parallel encoding
based on $\exists$-step plans and to generate a sequential encoding. Pasar uses
the \emph{grounding routine} deployed by the well-known planner \emph{Fast
  Downward}~\cite{FastDownward06} to translate planning tasks into a different
formalism and then encodes it to SAT using a parallel encoding.
% For planning tasks where this makespan cannot be determined with
% available computational resources, we use a \emph{sequential} encoding, where
% the minimal makespan equals the length of the shortest valid plan. Together with
% known bounds\footnote{Upper and lower bounds on plan length are available for
% 	some planning tasks from the \emph{Optimal track} that have \emph{unit cost}
% 	actions.} on the optimal plan length
%   we can generate satisfiable and unsatisfiable SAT formulas
%   for hard planning problems.
%\textcolor{red}{global note: should try to be consistent with present/past tense...}
The classical planning benchmarks were selected from the \emph{Satisfying} and
\emph{Optimal} tracks of the \emph{International Planning Competitions} 2014
\footnote{\url{https://helios.hud.ac.uk/scommv/IPC-14/repository/benchmarksV1.1.zip}}
and 2018 \footnote{\url{https://bitbucket.org/ipc2018-classical/domains}}. We
only selected planning domains with \emph{unit cost} and eliminated those that
% are designed to be challenging for the grounding routine and
take more than 100 GB of memory to encode into SAT. We ran both Pasar and
Madagascars $\exists$-configuration with a timeout of 3 hours on the
remaining instances to find the minimal makespans. For each planning task where
this process did not timeout, we generated a pair of satisfiable and
unsatisfiable SAT instances. A significant number of instances from this set were not used
as \solver{Minisat} could solve them in under $10$ minutes.
% From those we discarded the SAT formulas that \solver{Minisat} can solve in
% under $10$ minutes.
We augmented the remaining domains with the last unsatisfiable formulas
generated for planning task where the minimal makespan could not be found.
To generate the missing benchmarks we use a \emph{sequential} encoding
% , where the minimal makespan equals the length of the shortest valid plan
together with bounds\footnote{Upper and lower bounds on plan length are available for
	some planning tasks from the \emph{Optimal track} that have \emph{unit cost}
	actions.} on the optimal plan length.

In addition to the classical planning problems, we also included SAT instances
generated by \emph{Tree-REX} \cite{TreeRex19}, a planner for \emph{Hierarchical
	Task-networks} (HTN).
  In HTN planning additional domain knowledge besides the problem description is provided.
The HTN benchmarks were provided by
the solver author.

@@

Table \ref{tab:planningBenchmarkDist} shows the number of benchmarks generated
by each encoding.
For a complete list of the encoded planning tasks we refer to the
generation script\footnote{\url{https://satcompetition.github.io/2020/downloads/planning_generator.tar.xz}}.
The benchmarks of the Planning track adhere to the following naming convention:
${\langle \texttt{SAT/UNSAT} \rangle\_\langle \texttt{encoding} \rangle\_\langle
			\texttt{name} \rangle\_\langle \texttt{makespan}
			\rangle\text{.cnf}}$

@@

In this section we provide a high-level overview of the results of SAT Competition 2020.
%We focus on some interesting results.
Later on, we will provide overview of some of the key and new solving techniques implemented in best-performing solvers
(Section~\ref{sec:winners}) and as well as more in-depth analysis of the competition results (Section~\ref{sec:analysis}).

An overview of the top-10 solvers in each of the competition tracks discussed in the following
is provided in Table~\ref{tab:results}.

@@

Starting with the Main track, 
Figure~\ref{main:all} shows the cumulative solved instances plot of the best-performing solver of the strongest
10 teams (in short, the top-10 solvers)  together with the Virtual Best Solver (VBS---see also~\ref{subsec:vbs_explained}).
The best-performing  solver overall on the combination of satisfiable and unsatisfiable instances
is \solver{Kissat-sat} and the runner-up is \solver{Relaxed-newTech}. Notice that
\solver{Relaxed-newTech} solved more instances within the 2000-second limit.
Third place, based on the PAR-2 score, went to \solver{CMS-ccnr-lsids}. It solved two  instances less than
\solver{CaDiCaL-alluip-trail}, but on the other hand solved various formulas more quickly.
Similar observations have been made also in earlier recent SAT competitions where 
solvers were ranked based on the PAR-2 score.
Furthermore, we observe more differences in overall runtime performance among the top solvers than what has been observed
in the recent past competitions.

The four solvers \solver{Kissat-sat}, \solver{Relaxed-newTech}, \solver{CaDiCaL-alluip-trail}, and \solver{CMS-ccnr-lsids} 
performed significantly better that all other solvers submitted to the Main track (cf. Table~\ref{tab:results}).
A very interesting observation is that these four top solvers all have a different code base. This has not been observed for many years; more typically
many of the best-performing solver have been based on same code bases.

The majority of the overall performance differences
 between the top-4 solvers and the other solvers 
is due to performance differences on satisfiable instances;
see Figure~\ref{main:sat}. Indeed, in recent years, several techniques have been added to SAT solvers to improve
their performance on satisfiable instance. Examples of such techniques are the integration of a local search
solver and alternating between a SAT mode (infrequent restarts) and an UNSAT mode (frequent restarts
and variable-move-to-front~\cite{Ryan:2004:Thesis}). The best-performing solver in  Main SAT track on satisfiable instances is \solver{Relaxed-newTech}, followed by
\solver{Kissat-sat} and \solver{CMS-ccnr-lsids}.

@@

Overall, solvers performed much more similarly on
unsatisfiable instances than on satisfiable instances;
see Figure~\ref{main:unsat} for the runtime performance on unsatisfiable instances. 
Only \solver{Kissat-unsat}, the winner of the Main UNSAT track, 
performed significantly than all other participating solvers. 
It is therefore not surprising
that the VBS is reasonably close to \solver{Kissat-unsat}. %\solver{Kissat} is a new solver that has been implemented in C, while most solvers are implemented in C++. Further 
%details about \solver{Kissat} are given in Section~\ref{sec:kissat}.
The solvers \solver{CaDiCaL-trail}
and \solver{f2trc-s} placed, respectively, second and third in the Main UNSAT track.

@@

A further, more recent line of evolution in SAT solvers
is rooted in the \solver{CoMinisatPS} by Oh, which is itself a again a fork of \solver{Minisat}, introduced three-tier clause-management~\cite{Oh:2015:satunsat}. 
Building on \solver{CoMinisatPS}, 
The SAT solver \solver{Maple} appeared as a series of forks presenting innovative branching heuristics at SAT Competition 2016~\cite{MapleSAT}. 
The at-the-time award-winning variant \solver{MapleCOMSPS} by Liang et al.~implements a hybrid branching heuristic of classic variable-state independent decaying sum (VSIDS)~\cite{Zhang:2001:ClauseLearning} and the newer learning rate based branching (LRB) of \solver{CoMinisatPS}~\cite{Liang:2016:MapleCOMSPS}.

For SAT Competition 2017, Luo et al.~integrated learned clause minimization based on unit propagation (\solver{LCM}) in their award-winning \solver{Maple\_LCM\_Dist}~\cite{Luo:2017:LCM} which also uses the new branching heuristic \solver{Distance (Dist)} in an initial solving period~\cite{Xiao:2017:MapleLCMDist}. 
Using \solver{Maple\_LCM\_Dist}, Ryvchin and Nadel successfully integrated conditional chronological backtracking (\solver{ChronoBT})~\cite{Nadel:2018:CBT}, with
the introduction of the award-winning solver \solver{Maple\_LCM\_Dist\_ChronoBT} in SAT Competition~2018~\cite{Ryvchin:SC2018:MapleChronoBT}.

@@

Starting as a fork of \solver{Minisat} with the integration of special treatment for XOR constraints ~\cite{Soos:2009:Crypto},
\solver{CryptoMinisat} by Soos continues to be a state-of-the-art and feature-rich SAT solver. 
One highlight of \solver{CryptoMinisat} are its advanced data-logging capabilities for statistical analysis of SAT solver behavior~\cite{Soos:2019:ChrystalBall}.

Many independent and award-winning code-bases can be found among the SAT solvers written by Biere. 
The sequential SAT solver \solver{Lingeling} has been award-winning since SAT Competition~2011 and is still competitive in its parallel version \solver{Plingeling}~\cite{Biere:2012:Lingeling}. 
As of SAT Competition~2017, \solver{CaDiCaL} % \cite{Biere:SC2020} - or cite some older proceedings to match the "As of SAT Competition~2017" part?
by Biere is another independent representative of state-of-the-art SAT solvers.
Its improved re-implementation \solver{Kissat}~\cite{Biere:SC2020} was quite successful in the 2020 competition.

The sequential variant of the \emph{new code-base} \solver{ParaFrost} by Osama and Wijs was submitted for the first time to the 2020 SAT Competition,
 introducing several interesting new concepts~\cite{Osama:SC2020:Parafrost}.

@@

Sequential SAT solvers have been evaluated in the Main, Planning and Incremental Library track of SAT Competition~2020. 
$18$ teams submitted a total of $48$ solvers and configurations to the Main track and the Planning track of the competition, and $4$ solvers participated the Incremental Library track. 
%and were ranked by the number of application categories in which they won based on their PAR-2 score. 
%Submissions to the Incremental Library track have been evaluated with respect to $6$ different applications -- each with $50$ benchmark instances. 
%
Table~\ref{tab:solvers-main} displays an overview of the participating teams, base solvers and their variants. 
In the following, we provide a short overview of the best-performing solvers of 2020, based mainly on the solver descriptions submitted
to the 2020 competition proceedings by the authors of the individual solvers.

@@

Three configurations of \solver{Kissat} were submitted to the 2020  competition, including one default configuration and two specialized configurations which are specifically tailored towards satisfiable and unsatisfiable instances, respectively. 
\solver{Kissat} received \emph{four awards},achieving the first place in the Main track, 
the best score on unsatisfiable instances, the second-best score on satisfiable instances and the third place in the Planning track.

@@

Based on \solver{CaDiCaL}, its variants \solver{Trail} and \solver{AllUip} present implementations of a new \emph{Trail Saving} approach~\cite{Hickey:2020:TrailSaving} 
and an improved clause-learning \emph{Stable AllUIP}~\cite{Bacchus:SC2020}. 
Submitted were the three variants \solver{Trail}, \solver{AllUip} and \solver{AllUip+Trail}. 
The variants including \emph{Stable AllUIP} were the most successful in the 2020 competition, achieving in particular the  first place in the Planning track.

@@

\solver{Relaxed} integrates short runs of the local search solver \solver{CCAnr} through periodic export and import of assignments~\cite{Xindi:SC2019} and uses a probabilistic schedule for switching between $10$ phase selection modes. 
The \solver{Relaxed newTech} variant uses occurrence counts of variables in unsatisfied clauses during stochastic local search runs to recalculate variable priorities for their modified branching heuristic~\cite{Xindi:SC2020}.

@@

Six teams submitted a total of $10$ solvers and configurations to the Parallel track. 
In the following, we outline  the best-performing parallel solver implementations.

@@

\solver{Painless} uses a generic interface to integrate a solver and abstracts away the implementation details of parallelism and concurrent data-structures. Due to this,  implementations in \solver{Painless} boil down to implementing parallelization and clause sharing strategies~\cite{Frioux:2017:Painless}. 
\solver{Painless-MCOMSPS-STR} diversifies mainly via hard-coded configurations of the branching strategies LRB and VSIDS, and via sparse random initialization of variable polarities~\cite{Balyo:2015:HordeSATs}. 
Two special solver instances perform concurrent clause strengthening~\cite{Wieringa:2013:CCS} and Gaussian elimination, respectively. 
Regarding sharing, \solver{Painless-MCOMSPS-STR} uses an all-to-all strategy with a fixed-size clause buffer and an LBD filter with a dynamic limit~\cite{SC2020}.

@@

The parallel solver 
\solver{Plingeling} achieved the best score on unsatisfiable instances, the second place in the overall evaluation, and the third-best score on satisfiable instances. 
\solver{Plingeling} is built around the well-known \solver{Lingeling} and did not change since 2016. 
In a global master queue, \solver{Plingeling} shares unit clauses, equivalences and short clauses with a size limit of~$40$ and an LBD limit of~$8$. 
\solver{Plingeling} uses random seeds for diversification via variable polarities~\cite{Biere:SC2020,Biere:2012:Lingeling}.

@@

Five teams submitted a total of $6$ solvers and configurations in the Cloud track. 
In the following, we outline the best-performing massively parallel solver implementations.

@@

\solver{Mallob} is a fork of the  massively parallel SAT solver \solver{HordeSAT}~\cite{Balyo:2015:HordeSATs}. 
\solver{Mallob} performs dynamic load balancing through \emph{malleable job scheduling} in case the input contains several SAT instances of varying priority. 
This functionality is disabled in the submitted variant \solver{Mallob Mono}. 
\solver{Mallob} uses \solver{Lingeling-bcj} and as every 14th solver \solver{Mallob} spawns the stochastic local search solver \solver{YalSAT}~\cite{Biere:2018:LingelingYalsat}.
Diversification is done via randomized sparse initialization of branching scores. 
\solver{Mallob} shares clauses by organizing solvers in a binary tree in which clauses are asynchronously aggregated in a buffer which is passed along this tree from its leafs to the root. 
Each node performs a three-way merge of its local export buffer and the two incoming buffers. 
The aggregate that approaches the root of the binary tree is then broadcast to all solvers~\cite{Schreiber:SC2020}. 
\solver{Mallob} uses a global clause size limit and a dynamic size limit for the sharing buffer, which depends on its position in the binary tree and is larger the closer we get to the root. 
Clauses are sorted by their size during aggregation, such that smaller clauses are preferred over longer clauses. 
Duplicates are avoided by using a Bloom filter which is cleared periodically~\cite{Schreiber:SC2020}.

@@

\solver{TopoSAT~2}~\cite{Ehlers:2018:Toposat2} is a massively parallel SAT solver using \solver{Glucose~3}~\cite{Audemard:2018:Glucose}. 
The solver
 uses lock-free clause-exchange for solvers on the same machine and the message passing interface (MPI) to share clauses between machines~\cite{Hamadi:2009:ManySat}. 
\solver{TopoSAT~2} also strengthens clauses before export and delays clause import until the trail-size reaches a local minimum, as well as
diversifies via strategies used for branching, restarting, and clause forgetting~\cite{Ehlers:SC2020}.

@@

In this section, we provide additional analysis of the Main track results, going beyond the rankings.
In particular, we focus on metrics complementing the PAR-2 score used for ranking the solvers in the actual competition.

@@

The Virtual Best Solver (VBS) is a fictitious solver consisting 
of all the solvers that actually participated in the competition (or a specific track)
and an oracle which, when given an input instance,
invokes  the  solver which performed the best on that instance.
In this way, the performance of the VBS highlights a certain upper bound 
on performance achievable in principle by the 
 participating solvers (cf.~the figures in Section~\ref{sec:results}).

One can see that the VBS solves all the instances that
were solved by at least one solver and solves each instance in the best observed time.
By quantifying how much each participating solver contributes
to the performance of the VBS, we may attempt to establish which
technology (as represented by the solvers) is the most important
(and to what degree) in the observed state of the art in SAT solving.
We consider here the following three related metrics conceptually derived
from the notion of VBS.

@@

Table~\ref{tab:vbsMainALL} provides the result of applying the just-described three metrics 
to the full results of the Main track. We can see that the respective leaderboards 
are generally dominated by \solver{Kissat} in at least one of its configurations.
VBS-1 tells us that \solver{Kissat-unsat} was most often the fastest solver,
in particular on \SI{11.5}{\percent} of all the solved instances.

>>>>

For data-driven selection of benchmark instances, we used GBD~Tools\footnote{See \url{https://pypi.org/project/gbd-tools/}} which facilitates querying for instances with desired properties, e.g., by instance author, family, result or solver runtime~\cite{Iser:2018:GBD}. 
We also use GBD~Tools for distributing benchmark instances and their attributes to the general public.\footnote{See \url{https://gbd.iti.kit.edu}}

@@

The ``Bring Your Own Benchmarks'' (BYOB) rule, first  established in SAT Competition 2017~\cite{SC2017}, was again followed in 2020. 
By this rule, solver authors are required to submit 20 benchmark instances to accompany a solver submission in order to participate in the competition. 
These benchmarks have to be ``new'' in the sense that instances included in benchmark sets from previous SAT competitions are not allowed. 
Furthermore, at least ten of the required 20 instances are required to be ``interesting'', interpreted in loose terms by the standard \solver{Minisat} SAT solving needing at least one minute of runtime (on typical computing hardware) to solve an instance. 
It should be noted that new benchmarks could be submitted to the competition without needing to submit a solver.
As a results, as detailed in Table~\ref{tab:families}, 27 authors contributed a set of 1,260 new benchmark instances from a wide range of different instance families.

@@

To compile the set of 300 new instances, we first applied a hardness criterion by filtering out all instances solved by \solver{Minisat} in less than ten minutes.\footnote{We note that
the limit of ten minutes is again somewhat arbitrary.
This runtime hardness filter essentially aims to make sure that enough instances are included in the final benchmark set which allow for distinguishing in terms of relative performance between the best-performing competing solvers.
Unfortunately this limit was much greater that the requirements imposed for  ``interesting'' benchmark instances
by the BYOB rule. It could be more sensible to impose the same ten minutes limit also for an instance being ``interesting''. However, this would require more efforts
at least in terms of computation times from the solver authors in order to construct a set of interesting new benchmarks required by the BYOB rule.}
From the resulting 1,012 instances, in order to obtain a balanced benchmark set, 
we randomly selected $k$~instances per author using the value  $k$ which ensured that  the resulting set contains at least 300 instances.
This \emph{rule-based randomization} procedure is detailed as
Algorithm~\ref{algo:select}. Specifically,
we randomly selected seven satisfiable and seven unsatisfiable instances per author (Lines~3 and~4) and added instances of yet unknown result if this did not yield a total of 14~instances (Lines~5--7). 
Of the such obtained 308~instances, we randomly removed eight satisfiable instances, yielding a total of 114~satisfiable, 78~unsatisfiable and 108~instances of unknown satisfiability
status.

@@

For the Planning track,
the benchmark instances were generated using two SAT-based planners
\emph{Madagascar}~\cite{Madagascar14} and \emph{Pasar}~\cite{Pasar19}. We used
Madagascar both in its default configuration to generate a parallel encoding
based on $\exists$-step plans and to generate a sequential encoding. Pasar uses
the \emph{grounding routine} deployed by the well-known planner \emph{Fast
  Downward}~\cite{FastDownward06} to translate planning tasks into a different
formalism and then encodes it to SAT using a parallel encoding.
% For planning tasks where this makespan cannot be determined with
% available computational resources, we use a \emph{sequential} encoding, where
% the minimal makespan equals the length of the shortest valid plan. Together with
% known bounds\footnote{Upper and lower bounds on plan length are available for
% 	some planning tasks from the \emph{Optimal track} that have \emph{unit cost}
% 	actions.} on the optimal plan length
%   we can generate satisfiable and unsatisfiable SAT formulas
%   for hard planning problems.
%\textcolor{red}{global note: should try to be consistent with present/past tense...}
The classical planning benchmarks were selected from the \emph{Satisfying} and
\emph{Optimal} tracks of the \emph{International Planning Competitions} 2014
\footnote{\url{https://helios.hud.ac.uk/scommv/IPC-14/repository/benchmarksV1.1.zip}}
and 2018 \footnote{\url{https://bitbucket.org/ipc2018-classical/domains}}. We
only selected planning domains with \emph{unit cost} and eliminated those that
% are designed to be challenging for the grounding routine and
take more than 100 GB of memory to encode into SAT. We ran both Pasar and
Madagascars $\exists$-configuration with a timeout of three hours on the
remaining instances to find the minimal makespans. For each planning task where
this process did not timeout, we generated a pair of satisfiable and
unsatisfiable SAT instances. A significant number of instances from this set were not used
as \solver{Minisat} could solve them in under ten minutes.
% From those we discarded the SAT formulas that \solver{Minisat} can solve in
% under $10$ minutes.
We augmented the remaining domains with the last unsatisfiable formulas
generated for planning tasks where the minimal makespan could not be found.
To generate the missing benchmarks, we use a \emph{sequential} encoding
% , where the minimal makespan equals the length of the shortest valid plan
together with bounds\footnote{Upper and lower bounds on plan length are available for
	some planning tasks from the \emph{Optimal track} that have \emph{unit cost}
	actions.} on the optimal plan length.

In addition to the classical planning problems, we also included SAT instances
generated by \emph{Tree-REX} \cite{TreeRex19}, a planner for \emph{Hierarchical
	Task-networks} (HTN).
  In HTN planning, additional domain knowledge besides the problem description is provided.
The HTN benchmarks were provided by
the author of \emph{Tree-REX}.

@@

Table \ref{tab:planningBenchmarkDist} shows the number of benchmarks generated
by each encoding.
For a complete list of the encoded planning tasks we refer to the
generation script.\footnote{\url{https://satcompetition.github.io/2020/downloads/planning_generator.tar.xz}}
The benchmarks of the Planning track adhere to the following naming convention:
${\langle \texttt{SAT/UNSAT} \rangle\_\langle \texttt{encoding} \rangle\_\langle
			\texttt{name} \rangle\_\langle \texttt{makespan}
			\rangle\text{.cnf}}$

@@

In this section, we provide a high-level overview of the results of SAT Competition 2020.
%We focus on some interesting results.
Later on, we will provide an overview of some of the key and new solving techniques implemented in best-performing solvers
(Section~\ref{sec:winners}) as well as a more in-depth analysis of the competition results (Section~\ref{sec:analysis}).
An overview of the top-10 solvers in each of the competition tracks discussed in the following
is provided in Table~\ref{tab:results}.

@@

Starting with the Main track, 
Figure~\ref{main:all} shows the cumulative solved instances plot of the best-performing solver of the strongest
ten teams (in short, the top-10 solvers)  together with the Virtual Best Solver (VBS---see also~\ref{subsec:vbs_explained}).
The best-performing  solver overall on the combination of satisfiable and unsatisfiable instances
is \solver{Kissat-sat} and the runner-up is \solver{Relaxed-newTech}. Notice that
\solver{Relaxed-newTech} solved more instances within the 2000-second limit.
Third place, based on the PAR-2 score, went to \solver{CMS-ccnr-lsids}. It solved two  instances less than
\solver{CaDiCaL-alluip-trail}, but on the other hand solved various formulas more quickly.
Similar observations have been made also in earlier recent SAT competitions where 
solvers were ranked based on the PAR-2 score.
Furthermore, we observe more differences in overall runtime performance among the top solvers than what has been observed
in the recent past competitions.

The four solvers \solver{Kissat-sat}, \solver{Relaxed-newTech}, \solver{CaDiCaL-alluip-trail}, and \solver{CMS-ccnr-lsids} 
performed significantly better that all other solvers submitted to the Main track (cf. Table~\ref{tab:results}).
A very interesting observation is that these four top solvers all have a different code base. This has not been observed for many years; more typically
many of the best-performing solvers have been based on same code bases.

The majority of the overall performance differences
 between the top-4 solvers and the other solvers 
is due to performance differences on satisfiable instances;
see Figure~\ref{main:sat}. Indeed, in recent years, several techniques have been added to SAT solvers to improve
their performance on satisfiable instance. Examples of such techniques are the integration of a local search
solver and alternating between a SAT mode (infrequent restarts) and an UNSAT mode (frequent restarts
and variable-move-to-front~\cite{Ryan:2004:Thesis}). The best-performing solver in the Main SAT track on satisfiable instances is \solver{Relaxed-newTech}, followed by
\solver{Kissat-sat} and \solver{CMS-ccnr-lsids}.

@@

Overall, solvers performed much more similarly on
unsatisfiable instances than on satisfiable instances;
see Figure~\ref{main:unsat} for the runtime performance on unsatisfiable instances. 
Only \solver{Kissat-unsat}, the winner of the Main UNSAT track, 
performed significantly better than all other participating solvers. 
It is therefore not surprising
that the VBS is reasonably close to \solver{Kissat-unsat}. %\solver{Kissat} is a new solver that has been implemented in C, while most solvers are implemented in C++. Further 
%details about \solver{Kissat} are given in Section~\ref{sec:kissat}.
The solvers \solver{CaDiCaL-trail}
and \solver{f2trc-s} placed, respectively, second and third in the Main UNSAT track.

@@

A further, more recent line of evolution in SAT solvers
is rooted in the \solver{CoMinisatPS} by Oh, which is itself a again a fork of \solver{Minisat}, and which introduced three-tier clause-management~\cite{Oh:2015:satunsat}. 
Building on \solver{CoMinisatPS}, the SAT solver \solver{Maple} appeared as a series of forks presenting innovative branching heuristics at SAT Competition 2016~\cite{MapleSAT}. 
The at-the-time award-winning variant \solver{MapleCOMSPS} by Liang et al.~implements a hybrid branching heuristic of classic variable-state independent decaying sum (VSIDS)~\cite{Zhang:2001:ClauseLearning} and the newer learning rate based branching (LRB)~\cite{Liang:2016:MapleCOMSPS}.

For SAT Competition 2017, Luo et al.~integrated learned clause minimization based on unit propagation (\solver{LCM}) in their award-winning \solver{Maple\_LCM\_Dist}~\cite{Luo:2017:LCM} which also uses the new branching heuristic \solver{Distance (Dist)} in an initial solving period~\cite{Xiao:2017:MapleLCMDist}. 
In SAT Competition~2018, Ryvchin and Nadel successfully integrated conditional chronological backtracking (\solver{ChronoBT})~\cite{Nadel:2018:CBT} in their award award-winning solver \solver{Maple\_LCM\_Dist\_ChronoBT}~\cite{Ryvchin:SC2018:MapleChronoBT}.

@@

Also starting as a fork of \solver{Minisat} with the integration of special treatment for XOR constraints ~\cite{Soos:2009:Crypto},
\solver{CryptoMinisat} by Soos continues to be a state-of-the-art and feature-rich SAT solver. 
One highlight of \solver{CryptoMinisat} are its advanced data-logging capabilities for statistical analysis of SAT solver behavior~\cite{Soos:2019:ChrystalBall}.

Many independent and award-winning code-bases can be found among the SAT solvers written by Biere. 
The sequential SAT solver \solver{Lingeling} has been award-winning since SAT Competition~2011 and is still competitive in its parallel version \solver{Plingeling}~\cite{Biere:2012:Lingeling}. 
As of SAT Competition~2017, \solver{CaDiCaL} % \cite{Biere:SC2020} - or cite some older proceedings to match the "As of SAT Competition~2017" part?
by Biere is another independent representative of state-of-the-art SAT solvers and its improved re-implementation \solver{Kissat}~\cite{Biere:SC2020} was successful in the 2020 competition.

%The sequential variant of the \emph{new code-base} \solver{ParaFrost} by Osama and Wijs was submitted for the first time to the 2020 SAT Competition, introducing several interesting new concepts~\cite{Osama:SC2020:Parafrost}.

@@

Sequential SAT solvers have been evaluated in the Main, Planning and Incremental Library track of SAT Competition~2020. 
$18$ teams submitted a total of $48$ solvers and configurations to the Main track and the Planning track of the competition, and four solvers participated the Incremental Library track. 
%and were ranked by the number of application categories in which they won based on their PAR-2 score. 
%Submissions to the Incremental Library track have been evaluated with respect to $6$ different applications -- each with $50$ benchmark instances. 
%
Table~\ref{tab:solvers-main} displays an overview of the participating teams, base solvers and their variants. 
In the following, we provide a short overview of the best-performing solvers of 2020, based mainly on the solver descriptions submitted
to the 2020 competition proceedings by the authors of the individual solvers.

@@

Three configurations of \solver{Kissat} were submitted to the 2020  competition, including one default configuration and two specialized configurations which are specifically tailored towards satisfiable and unsatisfiable instances, respectively. 
\solver{Kissat} received \emph{four awards}, achieving the first place in the Main track, 
the best score on unsatisfiable instances, the second-best score on satisfiable instances and the third place in the Planning track.

@@

Based on \solver{CaDiCaL}, its variants \solver{Trail} and \solver{AllUip} present implementations of a new \emph{Trail Saving} approach~\cite{Hickey:2020:TrailSaving} 
and the improved clause-learning heuristic \emph{Stable AllUIP}~\cite{Bacchus:SC2020}. 
Submitted were the three variants \solver{Trail}, \solver{AllUip} and \solver{AllUip+Trail}. 
The variants including \emph{Stable AllUIP} were the most successful in the 2020 competition, achieving in particular the  first place in the Planning track.

@@

\solver{Relaxed} integrates short runs of the local search solver \solver{CCAnr} through periodic export and import of assignments~\cite{Xindi:SC2019} and uses a probabilistic schedule for switching between ten phase selection modes. 
The \solver{Relaxed newTech} variant uses occurrence counts of variables in unsatisfied clauses during stochastic local search runs to recalculate variable priorities for their modified branching heuristic~\cite{Xindi:SC2020}.

@@

Six teams submitted a total of ten solvers and configurations to the Parallel track. 
In the following, we outline  the best-performing parallel solver implementations.

@@

\solver{Painless} uses a generic interface to integrate a solver and abstracts away the implementation details of parallelism and concurrent data-structures. Due to this,  implementations in \solver{Painless} boil down to implementing parallelization and clause sharing strategies~\cite{Frioux:2017:Painless}. 
\solver{Painless-MCOMSPS-STR} diversifies mainly via hard-coded configurations of the branching strategies LRB and VSIDS, and via sparse random initialization of variable polarities~\cite{Balyo:2015:HordeSATs}. 
Two special solver instances perform concurrent clause strengthening~\cite{Wieringa:2013:CCS} and Gaussian elimination, respectively. 
Regarding sharing, \solver{Painless-MCOMSPS-STR} uses an all-to-all strategy with a fixed-size clause buffer and a dynamic LBD filter~\cite{SC2020}.

@@

The parallel solver 
\solver{Plingeling} achieved the best score on unsatisfiable instances, the second place in the overall evaluation, and the third-best score on satisfiable instances. 
\solver{Plingeling} is built around the well-known \solver{Lingeling} and did not change since 2016. 
In a global master queue, \solver{Plingeling} shares unit clauses, equivalences and short clauses with a size limit of~$40$ and an LBD limit of eight. 
\solver{Plingeling} uses random seeds for diversification via variable polarities~\cite{Biere:SC2020,Biere:2012:Lingeling}.

@@

Five teams submitted a total of six solvers and configurations in the Cloud track. 
In the following, we outline the best-performing massively parallel solver implementations.

@@

\solver{Mallob} is a fork of the  massively parallel SAT solver \solver{HordeSAT}~\cite{Balyo:2015:HordeSATs}. 
\solver{Mallob} performs dynamic load balancing through \emph{malleable job scheduling} in case the input contains several SAT instances of varying priority. 
This functionality is disabled in the submitted variant \solver{Mallob Mono}. 
\solver{Mallob} uses \solver{Lingeling-bcj} and as every 14th solver \solver{Mallob} spawns the stochastic local search solver \solver{YalSAT}~\cite{Biere:2018:LingelingYalsat}.
Diversification is done via randomized sparse initialization of branching scores. 
\solver{Mallob} shares clauses by organizing solvers in a binary tree in which clauses are asynchronously aggregated in a buffer which is passed along this tree from its leafs to the root. 
Each node performs a three-way merge of its local export buffer and the two incoming buffers. 
The aggregate that approaches the root of the binary tree is then broadcast to all solvers~\cite{Schreiber:SC2020}. 
\solver{Mallob} uses a global clause size limit and a dynamic size limit for the sharing buffer, which depends on its position in the binary tree and is larger the closer we get to the root. 
Clauses are sorted by their size during aggregation, such that smaller clauses are preferred over longer clauses. 
Duplicates are avoided by using a Bloom filter which is cleared periodically~\cite{Schreiber:2021:Mallob}.

@@

\solver{TopoSAT~2}~\cite{Ehlers:2018:Toposat2} is a massively parallel SAT solver using \solver{Glucose~3}~\cite{Audemard:2018:Glucose}. 
The solver
 uses lock-free clause-exchange for solvers on the same machine and the message passing interface (MPI) to share clauses between machines~\cite{Hamadi:2009:ManySat}. 
\solver{TopoSAT~2} strengthens clauses before export and delays clause import until the trail-size reaches a local minimum. 
\solver{TopoSAT~2} diversifies via strategies used for branching, restarting, and clause forgetting~\cite{Ehlers:SC2020}.

@@

In this section, we provide an additional analysis of the Main track results, going beyond the rankings.
In particular, we focus on metrics complementing the PAR-2 score used for ranking the solvers in the actual competition.

@@

The Virtual Best Solver (VBS) is a fictitious solver consisting 
of all solvers that actually participated in the competition (or a specific track)
and an oracle which, when given an input instance,
invokes  the  solver which performed the best on that instance.
This way, the performance of the VBS highlights a certain upper bound 
on the performance achievable in principle by the 
 participating solvers (cf.~the figures in Section~\ref{sec:results}).

One can see that the VBS solves all instances that
were solved by at least one solver and solves each instance in the best observed time.
By quantifying how much each participating solver contributes
to the performance of the VBS, we may attempt to establish which
technology (as represented by the solvers) is the most important
(and to what degree) in the observed state of the art in SAT solving.
We consider here the following three related metrics conceptually derived
from the notion of VBS.

@@

Table~\ref{tab:vbsMainALL} provides the result of applying the just-described three metrics 
to the full results of the Main track. We can see that the respective leaderboards 
are generally dominated by \solver{Kissat} in at least one of its configurations.
VBS-1 tells us that \solver{Kissat-unsat} was most often the fastest solver,
in particular on \SI{11.5}{\percent} of the solved instances.
